---
title: "p8105_hw2_cm4070"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(readxl)

```

## Problem 1

### Read and clean Mr. Trash Wheel dataset

```{r, trashwheel_problem1}
mrtrashwheel_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 1, range = "A2:N535") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
  sports_balls = as.integer(sports_balls)) 
```

### Read and clean Precipitation datasets

```{r, precipitation18_problem1}
precipitation2018_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 7, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2018")
```

```{r, precipitation19_problem1}
precipitation2019_data =
  read_excel("./data/Trash-Wheel-Data.xlsx", 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2019")
```

### Combine 2018 and 2019 Precipitation datasets 

```{r, precipitation_problem1}
precipitation_data = 
  bind_rows(precipitation2018_data, precipitation2019_data) %>%
  mutate(
    month = month.name[month]
  )
```

### Describe these data 

```{r, echo = FALSE}
year2019_data = filter(mrtrashwheel_data, year == "2019")
```

There are `r nrow(mrtrashwheel_data)` observations  in `mrtrashwheel_data` and key variables are `r names(mrtrashwheel_data)`. There are `r nrow(precipitation_data)` observations in `precipitation_data` and key variables are `r names(precipitation_data)`. The total precipitation in 2018 was `r sum(pull(precipitation2018_data, total))` inches. The median sports balls in 2019 was `r median(pull(year2019_data, sports_balls))`.

## Problem 2 

### Read and Clean pols-month dataset

```{r, pols_problem2, warning = FALSE}
pols_data = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  unite('president', prez_dem,prez_gop, sep = "") %>%
  mutate(president = as.numeric(president)) %>%
  mutate(president = recode(president, '10' = 'dem', '01' = 'gop')) %>%
  mutate(month = str_to_lower(month)) %>%
  select(-day)
```

### Read and Clean snp dataset

```{r, snp_problem2}
snp_data = 
  read_csv(file = "./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(month = str_to_lower(month)) %>%
  select(-day)
```

### Read and Clean unemployment dataset

```{r, unemployment_problem2}
unemployment_data = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec, 
    names_to = "month", 
    values_to = "unemployment")
```
