---
title: "p8105_hw2_cm4070"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(readxl)

```

## Problem 1

### Read and clean `Mr. Trash Wheel` dataset

```{r, trashwheel_problem1}
mrtrashwheel_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 1, range = "A2:N535") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
  sports_balls = as.integer(sports_balls)) 
```

### Read and clean `Precipitation` datasets

```{r, precipitation18_problem1}
precipitation2018_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 7, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2018")
```

```{r, precipitation19_problem1}
precipitation2019_data =
  read_excel("./data/Trash-Wheel-Data.xlsx", 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2019")
```

### Combine 2018 and 2019 `Precipitation` datasets 

```{r, precipitation_problem1}
precipitation_data = 
  bind_rows(precipitation2018_data, precipitation2019_data) %>%
  mutate(
    month = month.name[month]
  )
```

### Describe these data 

```{r, echo = FALSE}
year2019_data = filter(mrtrashwheel_data, year == "2019")
```

There are `r nrow(mrtrashwheel_data)` observations  in `mrtrashwheel_data` and key variables are `r names(mrtrashwheel_data)`. There are `r nrow(precipitation_data)` observations in `precipitation_data` and key variables are `r names(precipitation_data)`. The total precipitation in 2018 was `r sum(pull(precipitation2018_data, total))` inches. The median sports balls in 2019 was `r median(pull(year2019_data, sports_balls))`.

## Problem 2 

### Read and Clean `pols-month` dataset

```{r, pols_problem2, warning = FALSE}
pols_data = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  unite('president', prez_dem,prez_gop, sep = "") %>%
  mutate(president = as.numeric(president)) %>%
  mutate(president = recode(president, '10' = 'dem', '01' = 'gop')) %>%
  mutate(month = str_to_lower(month)) %>%
  select(-day)
```

### Read and Clean `snp` dataset

```{r, snp_problem2}
snp_data = 
  read_csv(file = "./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(month = str_to_lower(month)) %>%
  select(-day)
```

### Read and Clean `unemployment` dataset

```{r, unemployment_problem2}
unemployment_data = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec, 
    names_to = "month", 
    values_to = "unemployment") %>%
  mutate(year = as.character(year))
```

### Join `snp` into `pols` datasets then `unemployment`

```{r, joindataset_problem2}
snppols_data = 
  left_join(pols_data, snp_data, by = "year")

conjoinedunemployment_data = 
  left_join(unemployment_data, snppols_data, by = "year")
```

### Describe these data 

The `pols_data` dataset contained poll data from `r nrow (pols_data)` observations and `r names(pols_data)` variables. The `snp_data` dataset contained `r nrow(snp_data)` observations and `r names(snp_data)` variables. 
The `unemployment_data` dataset contained `r nrow(unemployment_data)` observations and `r names(unemployment_data)` variables. 
The `conjoinedunemployment_data` dataset contained
`r nrow(conjoinedunemployment_data)` observations and `r names(conjoinedunemployment_data)` variables. Therefore it included all variables within each dataset, but combined year through joining.

## Problem 3 

### Load and tidy `popular_baby_names` dataset

```{r, popular_baby_names_prolem3}
babynames_data = read.csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    child_s_first_name = 
    str_to_lower(child_s_first_name)) %>% 
  mutate(gender = str_to_lower(gender)) %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>% 
  relocate(child_s_first_name) %>%
  distinct(.keep_all = TRUE)
```

### Make a table for "Olivia" name

```{r, olivia_table_problem3}
olivia_table = 
  filter(babynames_data, child_s_first_name == "olivia", gender == "female") %>%
  select(-child_s_first_name, -gender, -count) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank") %>%
  relocate(
    "ethnicity", "2011", "2012", "2013", "2014", 
    "2015", "2016")
```

## Make a table for Male's most popular name

```{r, male_table_problem3, warning= FALSE}
male_table = 
  filter(babynames_data, gender == "male") %>%
  select(-gender, -count) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "child_s_first_name") %>%
  relocate(
    "ethnicity", "rank", "2011", "2012", "2013", 
    "2014", "2015", "2016") 
```

## Make a plot

_This code creates a scatterplot showing the number of male, non hispanic children in 2019 with a name againist the popularity of that name._

```{r ggplot, warning=FALSE}
subset_babynames = filter(babynames_data, gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016")

ggplot(subset_babynames, aes(x = rank, y = count)) + geom_point()
```

