---
title: "p8105_hw2_cm4070"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(readxl)
```

## Problem 1

### Read and clean Mr. Trash Wheel dataset

```{r, trashwheel_problem1}
mrtrashwheel_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 1, range = "A2:N535") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
  sports_balls = as.integer(sports_balls)) 
```

### Read and clean Precipitation datasets

```{r, precipitation18_problem1}
precipitation2018_data = 
  read_excel("./data/Trash-Wheel-Data.xlsx", 7, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2018")
```

```{r, precipitation19_problem1}
precipitation2019_data =
  read_excel("./data/Trash-Wheel-Data.xlsx", 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = "2019")
```

### Combine 2018 and 2019 Precipitation datasets 

```{r, precipitation_problem1}
precipitation_data = 
  bind_rows(precipitation2018_data, precipitation2019_data) %>%
  mutate(
    month = month.name[month]
  )
```

### Describe these data 

```{r, echo = FALSE}
year2019_data = filter(mrtrashwheel_data, year == "2019")
```

There are `r nrow(mrtrashwheel_data)` observations  in `mrtrashwheel_data` and key variables are `r names(mrtrashwheel_data)`. There are `r nrow(precipitation_data)` observations in `precipitation_data` and key variables are `r names(precipitation_data)`. The total precipitation in 2018 was `r sum(pull(precipitation2018_data, total))` inches. The median sports balls in 2019 was `r median(pull(year2019_data, sports_balls))`.




